---
title: "Data_Exploration_MarkDown"
author: "Kristen Higashi"
date: "08/07/2022"
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/krist/Documents/SU_OMSBA/Summer_2022/DataExplorationAssignment/data")

```

## Data Exploration Assignment

### RESEARCH

The College Scorecard was released at the start of September 2015. Among colleges that predominantly grant bachelor’s degrees, did the release of the Scorecard shift student interest to high-earnings colleges relative to low-earnings ones (as proxied by Google searches for keywords associated with those colleges)?

This study seeks to determine how the College Scorecard release (starting in 2015) affected student interest by Google search keywords.

Share of students earning over $25,000/year (threshold earnings) 6 years after entry
### Notes on Data 

However, college interest may be swayed by many factors, including potential student demographics (race and family legacy), degree interest, and  standing() but the data available for this study only included:

1. College Historical affiliations, location, and attendee demographics
retention. 

(including physical geographical location or if the school is virtually inteded). IE 
2. Price of attending and potential debt accrual --- including Pell grants
3. Study opportunities / Available degrees
4. Retention (as a measure of approval)




#### Load Libraries
```{r}

library(car)
library(vtable)
library(jtools)
library(vtable)
library(purrr)
library(tidyverse)
library(dplyr)
library(lubridate)

# Set Working Directory
setwd("C:/Users/krist/Documents/SU_OMSBA/Summer_2022/DataExplorationAssignment/data")
```

## Load/Join Data Tables
```{r}
score_card <- read.csv("Most+Recent+Cohorts+(Scorecard+Elements).csv")

# Bind "trends_up_to_" data
trends_all <- data.frame()
for (data in list.files(pattern = "trends_up_to_")){
  temp <- read.csv(data)
  trends_all <- rbind(trends_all, temp)
}

# Scorecard and ID CSVs
score_card <- read.csv("Most+Recent+Cohorts+(Scorecard+Elements).csv")
id_name <- read.csv("id_name_link.csv")%>%
  distinct(schname, .keep_all = TRUE)
coll_ids <- left_join(id_name, score_card, by = c("unitid"="UNITID"))
id_name <- read.csv("id_name_link.csv")


```


```{r}
# Full join - Removes all "not currently open institutions"
## College information
df_undg <- 
  # Remove duplicate "schname" and join tables
  unique(left_join(trends_all, coll_ids, by = c("schname" = "schname"))) %>%
  
  # filter for schools that predominantly grant bachelor degrees AND currently operating
  filter(PREDDEG == 3
         , !is.null(monthorweek)
         , CURROPER == 1)%>%
  # earnings
  mutate(num_earnings= as.numeric(md_earn_wne_p10.REPORTED.EARNINGS)) %>%
  filter(!is.na(num_earnings))%>%
  
  # Convert times
  mutate(month  = floor_date(ymd(str_sub(monthorweek, 0,10)), unit = c("month"))
         ,  yr = as.Date(str_sub(monthorweek, 1, 4), "%Y")) %>%
  mutate(pre_scorecard = as.factor(month < ymd("2015-01-01")))%>%
  mutate(key = paste(opeid,month, sep = "_"), date_state = paste(STABBR, monthorweek, sep="_"))%>%


  # Mutate variables to correct class(es)
  select(key, date_state, unitid, opeid, opeid6,	schname, INSTNM, STABBR
         , month, yr, monthorweek, pre_scorecard
         , STABBR, num_earnings, keyword, keynum, index
         , c(HBCU:WOMENONLY), RELAFFIL, DISTANCEONLY
         , starts_with("PCIP")
         , starts_with("UGDS")
         , ACTCMMID, SAT_AVG, SAT_AVG_ALL, starts_with("NPT")
         , starts_with("RET")
         , starts_with("GRAD")
         , RPY_3YR_RT_SUPP, PCTFLOAN, PCTPELL, UG25abv, C150_4_POOLED_SUPP.REPORTED.GRAD.RATE, PPTUG_EF)%>%
  mutate_at(vars(c(HBCU:WOMENONLY), RELAFFIL, DISTANCEONLY), as.factor)%>%
  mutate_at(vars(matches("PCIP"), matches("UGDS")), as.numeric)%>%
  mutate_at(vars(matches("NPT"), matches("RET"), matches("GRAD"), RPY_3YR_RT_SUPP, PCTFLOAN, PCTPELL), as.numeric)%>%
  mutate(UG25abv = as.numeric(UG25abv), C150_4_POOLED_SUPP.REPORTED.GRAD.RATE= as.numeric( C150_4_POOLED_SUPP.REPORTED.GRAD.RATE), PPTUG_EF = as.numeric(PPTUG_EF))%>%
  mutate_at(vars(ACTCMMID, SAT_AVG, SAT_AVG_ALL), as.numeric)%>%

  # Determine count of searches and the max number of keywords searched
  group_by(schname, month)%>%
  mutate(ct_index = sum(index), num_keywords = max(keynum))%>%
  ungroup()%>%
  filter(!is.na(month), !is.na(num_earnings))%>%

  # Determine average, standard deviation, and avg+sd for each key word
  group_by(schname, month, keynum)%>%
  mutate(limit_index = mean(index) + sd(index))%>%
  ungroup()%>%
  
  # Determine if index is deemed above average (ie above avg+sd in state). TRUE/FALSE (True if larger than avg+stdev)
  mutate(exc_index = index >= limit_index)%>%
  filter(!is.na(exc_index)) %>% 

  # Determine average, standard deviation, and avg+sd for earning (by state and date window since states will often have different average wages)
  group_by(date_state)%>%
  mutate(limit_earnings = mean(num_earnings) + sd(num_earnings))%>%
  ungroup()%>%

  # Determine if earnings are deemed above average (ie above avg+sd in state). TRUE/FALSE (True if larger than avg+stdev)
  mutate(exc_earning = num_earnings>= limit_earnings)%>%
  filter(!is.na(exc_earning))

```


## Data Exploration
```{r}
summary(df_undg)



```
# 
# ```{r}
# 
# 
# earns <- lm(num_earnings~UGDS+ UGDS_WHITE+ UGDS_BLACK+ UGDS_HISP+ UGDS_ASIAN+ UGDS_AIAN+ UGDS_NHPI+ UGDS_2MOR+ UGDS_NRA+ UGDS_UNKN, data = df_undg) %>%
#   export_summs(error_format = 't = {statistic}, p = {p.value}', digits=5)
# earns
# 
# dem <- df_undg%>%
#  (c(HBCU:WOMENONLY), RELAFFIL, DISTANCEONLY)
# mutate(tot=sum(c(HBCU:WOMENONLY)))
# 
#  dem_lm <- lm(ct_index ~ num_earnings, data = df_undg) 
# 
# 
#  # Mark as private v public institution -->
# 
#   ugds_npt <- df_undg%>% 
# select(num_earnings, exc_earning, starts_with("UGDS"), starts_with("NPT"))%>%
# 
# 
# 
# 
# lm(is_pub~PCIP01 + PCIP03 + PCIP04 + PCIP05 + PCIP09 + PCIP10 + PCIP11 + PCIP12 + PCIP13 + PCIP14 + PCIP15 + PCIP16 + PCIP19 + PCIP22  -->
#     + PCIP23 + PCIP24 + PCIP25 + PCIP26 + PCIP27 + PCIP29 + PCIP30 + PCIP31 + PCIP38 + PCIP39 + PCIP40 + PCIP41 + PCIP42 + PCIP43 + PCIP44 + PCIP45 + PCIP46 + PCIP47 + PCIP48 + PCIP49 + PCIP50 + PCIP51 + PCIP52 + PCIP54, data = df_undg)%>% 
#    export_summs(error_format = 't = {statistic}, p = {p.value}', digits=5)
# 
# 
# 
# lm_dem_current <- lm(ct_index~is_pub + UGDS+ UGDS_WHITE+ UGDS_BLACK+ UGDS_HISP+ UGDS_ASIAN+ UGDS_AIAN+ UGDS_NHPI+ UGDS_2MOR+ UGDS_NRA+ UGDS_UNKN, data = df_undg)
# 
# lm_dem_flag <- lm(ct_index~ HBCU + PBI + ANNHI + TRIBAL + AANAPII + HSI + NANTI + MENONLY + WOMENONLY + DISTANCEONLY, data = df_undg)
# 
# lm_earnings <-  lm(ct_index~num_earnings+ exc_earning, data = df_undg)
# 
# export_summs(lm_dem_current, lm_dem_flag, lm_earnings)
# 
# linearHypothesis(lm_dem_current)
# #  -->
# #  -->
# # sumpcip <- df_undg%>% -->
# #   select(key, starts_with ("PCIP"))%>% -->
# #   replace(is.na(.),0)%>% -->
# #   transmute(key, sum = rowSums(across(where (is.numeric)))) -->
# #  -->
# 
# # Regression between the total earnings based on  -->
# 
# df_undg
# 
# 
# 
# 
# # #  -->
# # # # Merge keyw_ct and avg_earnings -->
# # # key_earnings <- avg_earnings%>% -->
# # #   left_join(keyw_ct, by= c("monthorweek", "STABBR"), copy = FALSE )%>% -->
# # #   drop_na()%>% -->
# # #   mutate(high_earn = limit_earnings<num_earnings)%>% -->
# # #   select(unitid, schname, monthorweek, limit_earnings, keynum, index, ct_index, num_keywords, high_earn) -->
# # #  -->
# # #  -->
# # # #  -->
# # # #  -->
# # # # ## Regress key earnings -->
# # # # reg_keyword_earnings <- lm(ct_index ~ high_earn + num_keywords, data = key_earnings) -->
# # # # export_summs(reg_keyword_earnings) -->
# # #  -->
# # #  -->
# # # #  -->
# # # # test_ <- lm(ct_index) -->
# # # # regression <- lm(ct_index~ excess + , data = merged) -->
# # # # export_summs(regression) -->
# # # #  -->
# # #  -->
# # #  -->
# # #  -->
# # #  -->
